Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-29 paper/paper.tex
updated references

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[An atypical \lya\ dwarf galaxy]{
Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy}
\author[J.E. Forero-Romero et al.]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Max Gronke$^2$, 
Maria Camila Remolina-Guti\'errez$^1$,
\newauthor
Nicol\'as Garavito-Camargo$^3$, 
Mark Dijkstra$^2$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
$^2$ Institute of Theoretical Astrophysics, University of Oslo,
Postboks 1029 Blindern, NO-0315 Oslo, Norway.\\
$^3$ Department of Astronomy, University of Arizona, 933 North Cherry
Avenue, Tucson, AZ 85721, USA. 
}

@@

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the multiphase model has more chances to be correct.
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$6\times 10^{11}$ M$_{\odot}$.
If future kinematic maps of \tol\ confirm the velocities suggested by the
multiphase model, it would provide additional support to expect this
condition to be present  in primordial \lya\ emitting galaxies.  
\end{abstract}

@@

Primordial galaxies have not been detected yet. 
However, dwarf star forming galaxies with a low metallicity content
are seen as templates to understand the early galaxy evolution process. 
Over fifty years ago it was realized that young galaxies could be
detected through a strong \lya\ line emission \citep{PartridgePeebles}.

This theoretical prediction was only confirmed thirty years later on
distant, relatively young, not primordial, galaxies \citep{1998ApJ...498L..93D}.
Currently Lyman Alpha Emitting (LAE) galaxies are commonly targeted
in surveys. 
The presence of the \lya\ emission line provides confirmation of
the distance to a galaxy while provides clues about the stellar
population and inter-stellar medium conditions regulating the
\lya\ emission.
A careful clustering analysis of LAEs can also yield clues about their link
to dark matter halos
\citep{2004AJ....128.2073H,2007ApJ...671..278G,2007ApJ...668...15K,2008MNRAS.391.1589O,2010MNRAS.409..184P,2013MNRAS.431.1777G,2016ApJ...828....5M}.

The \lya\ emission line is not exclusive of distant galaxies. 
There are local Universe surveys that target \lya\ emission in nearby
dwarf star forming galaxies.
The study of nearby LAE samples has allowed the study of other
indicators that might be more difficult to obtain for distant galaxies
such as galaxy morphology, dust attenuation, neutral hydrogen contents and
ionization state. See \cite{Hayes15} and references therein for a review.

However, the physical interpretation of \lya\ observations is
not straightforward \citep{LARS,2015ApJ...805...14R}. 
This is due to the resonant nature of the \lya\ line. 
A \lya\ photon follows a diffusion-like process before escaping
the galaxy or being absorbed by dust. 
The resulting line profile becomes sensitive to the dynamical, chemical
and thermal conditions in the interstellar medium. 
There are few analytical tools available to interpret the
emerging \lya\ line
\citep{Harrington73,1991ApJ...370L..85N,LoebRybicki,2006ApJ...645..792T}. 
They are applicable only in cases of highly symmetrical
conditions of the gaseous hydrogen distribution in the galaxy model,
which are hardly met in real astrophysical systems. 
For these reasons the interpretation of \lya\ observations
requires state-of-the-art Monte Carlo radiative transfer simulations.

Observed \lya\ line profiles usually present a single peak shifted
redwards from the line's center. 
Sometimes a double peak is present but the asymmetry persists with 
the peak on the red side being stronger. 
This can be explained as the results of multiple \lya\ photon
scatterings through an outflowing shell of neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

\tol\ is a compact star forming dwarf galaxy that presents a
strong \lya\ emission with puzzling 
features: the line is highly symmetric, single peaked and broad 
\citep{Thuan97}.
This kind of profile also raises the question whether some high
redshift LAEs have asymmetric lines because the blue half was
truncated by the intergalactic medium \citep{2007MNRAS.377.1175D}. 
In this case the \lya\ radiation could emerge as a low surface
brightness glow, which may be connected to \lya\ halos, while also
influencing the way LAEs can be used as a probe of reionization,
see the review by \cite{2014PASA...31...40D} and the references
therein.

Attempts to explain the atypical \lya\ features in \tol\ with
conventional models based on a expanding shell have not been succesful
so far \citep{mashesse03,2015A&A...578A...7V}. 
This justifies a new study using two new physically motivated
situations: a multiphase outflowing interstellar medium and bulk rotation for a
galaxy without outflows.

The main motivation for the outflowing multiphase model (as presented
in \cite{Gronke2016}) is that some dwarf galaxies are expected to
present outflows.  
Observationally, outflows have been detected in few local dwarf galaxies
\citep{1998ApJ...506..222M,2005MNRAS.358.1453O}. 
Besides, clumpy multiphase outflows are capable to explain \lya\ features around
star-forming galaxies
\citep{2010ApJ...717..289S,2012MNRAS.424.1672D}.

@@

The models we use correspond to simplified geometrical configurations.
This allows us to perform a deep exploration of parameter space and
gain some physical insight into \tol's kinematic properties.
In this paper we show, for the first time in the literature, that \tol's
\lya\ profile can be explicitly modeled be either of these two
models.

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be sucessful.
We focus on presenting two simple and succesful models that
manage, for the first time, to account for \tol's \lya\ line main
features.

In the next section We review the observational characteristics of
\tol, then we summarize the main features in the multiphase and
rotation models to explain how we fit their free parameters 
to the \tol's \lya\ line shape.
We use the results to interpret them in terms of the galaxy's
dynamical mass and argue why the multiphase model should be prefered
over the rotation model.
We close by discussing the additional observational information needed
to clarify the kinematic nature of \tol.

@@

\begin{table*}
  \centering
\begin{tabular}{l|lrrl}
\hline
  Parameter  & Description & Fiducial value & Allowed range & Units \\ \hline\hline
$v_{\infty,\,{\rm cl}}$ & Radial cloud velocity & $100.0$ & [$0.0$, $800.0$] & $\,{\rm km\,s}^{-1}$ \\
$\sigma_{\rm {\rm cl}}$ & Random cloud motion &$40.0$ & [$5.0$, $100.0$] & $\,{\rm km\,s}^{-1}$ \\
$P_{\rm {\rm cl}}$ & Probability to be emitted in cloud & $0.35$ & [$0.0$, $1.0$] &  \\
$r_{\rm {\rm cl}}$ & Cloud radius & $100.0$ & [$30.0$, $200.0$] & $\,{\rm pc}$ \\
$H_{\rm em}$ & Emission scale radius & $1000.0$ & [$500.0$, $3.0\times 10^{3}$] & $\,{\rm pc}$ \\
$f_{\rm {\rm cl}}$ & Cloud covering factor & $3.5$ & [$0.8$, $8.0$] &  \\
$T_{\rm {\rm ICM}}$${}^{\dagger}$ & Temperature of ICM & $10^{6}$ & [$3.0\times 10^{5}$, $5.0\times 10^{7}$] & $\,{\rm K}$ \\
$n_{\rm HI,\,{\rm ICM}}$${}^{\dagger}$ & \HI number density in ICM & $5.0\times 10^{-8}$ & [$10^{-12}$, $10^{-6}$] & $\,{\rm cm}^{-3}$ \\
$\sigma_{\rm i}$ & Width of emission profile & $50.0$ & [$5.0$, $100.0$] & $\,{\rm km\,s}^{-1}$ \\
$T_{\rm {\rm cl}}$${}^{\dagger}$ & Temperature in clouds & $10^{4}$ & [$5.0\times 10^{3}$, $5.0\times 10^{4}$] & $\,{\rm K}$ \\
$\beta_{\rm cl}$ & Steepness of the radial velocity profile & $1.5$ & [$1.1$, $2.5$] &  \\
$\tilde\sigma_{\rm d, cl}$${}^{\dagger}$ & Dust content in clumps &
  $3.2\times10^{-22}$ & [$4.7\times 10^{-24}$, $1.6\times 10^{-21}$] &
  $\,\cm^{2}$ \\
$\zeta_d$${}^{\dagger}$ & Ratio of ICM to cloud dust abundance &
  $0.01$ & [$10^{-4}$, $0.1$] & \\
$n_{\rm HI,\,{\rm cl}}$${}^{\dagger}$ & \HI number density in
  clouds & $0.35$ & [$0.03$, $3.0$] & $\,{\rm cm}^{-3}$ \\
\hline
\end{tabular}
  \caption{Overview of the parameters in the multiphase model and its fiducial
    values. Variables marked with ${}^{\dagger}$ were drawn in
    log-space. Details of the implementation are in \citet{Gronke2016}.}
  \label{tab:models}
\end{table*}

@@

Archival Chandra X-ray data \footnote{\url{http://cxc.harvard.edu/cda/}} does not show any
detection for \tol. 
This lack of detection motivates our assumption that \lya\ emission is
powered by star formation only.

The observed flux for the \lya\ line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \citep{Thuan97}.
The \lya\ Equivalent Width is $70$\AA\ and its H$\beta$ flux is 
$1.62\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \AA${-1}$
which gives a \lya/H$\beta$ flux ratio of
4.9$\pm$0.1 \citep{Izotov04}.
Comparing the \lya/H$\beta$ ratio with the theoretical
expectation from case B recombination of $23.3$ \citep{Hummer1987} one
can estimate an escape fraction of $20$\% for \lya\ radiation.

The \lya\ flux values correspond to a luminosity of
$L_{Ly\alpha}=2.2\times 10^{42}$ erg s$^{-1}$, which in turn
translates  into a lower bound for the star formation rate of $2.0$
M$_{\odot}$ yr$^{-1}$ after using a standard conversion factor between
luminosity and star formation rate of $9.1\times 10^{-43}$
$L_{Ly\alpha}$ M$_{\odot}$ yr$^{-1}$ 
without any
correction by extinction
\citep{Kennicutt98}.

@@

We computed the projected half-light radius to be $R_s=1.5\pm0.1$ kpc 
from the surface brightness profiles reported by \cite{2003A&A...410..481N}. 
Assuming spherical geometry, one can translate this value into a 3D
half-light radius of $r_s=3R_s/2=2.25$ kpc.
Imaging observations by \cite{2001AJ....121..169F} show that \tol\ is
an isolated field galaxy not belonging to a group or cluster.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.55\textwidth]{CLARA-TOL-main.pdf}
\caption{Broad, single peaked and highly symmetric \lya\ emission of \tol.
  Dots correspond to the observational data \citep{mashesse03}. 
The thick red and thin blue curves represent our best fit
to the data using the full radiative transfer simulation with a
rotation and multiphase model, respectively. 
These two different models are able to reproduce the main morphological
features of \tol's \lya\ line emission.\label{fig:results}}
\end{center}
\end{figure*}

@@

The rotation model corresponds to the work presented in
\citep{GaravitoCamargo2014} based on the Monte Carlo code
\texttt{CLARA} \citep{CLARA}. 
In that model the \lya\ photons are propagated 
within a spherical and homogeneous cloud of HI gas undergoing solid
body rotation.
The sphere is fully characterized by three parameters: the HI line's
center optical  depth $\tau$ measured from the center to its surface, the HI
temperature $T$, and the linear surface velocity $V_{\rm max}$.  
The observed line profile also depends on $\theta$,  the angle between the plane
perpendicular to the rotation axis and the observational
line-of-sight.  
In this model, dust only changes the overall line
normalization and only weakly its shape, i.e. 
dust cannot change the line symmetry or induce a change in the number
of line peaks, moreover it does not change the line width by more than
$1\%$ ($5$ \kms\ in the case of \tol), an effect too small to be
observed at the resolution at which we have \tol's line profile and
also negligeable compared to the influence of the other free parameters in the model.
For these reasons we do not include any dust model.

We use an analytical approximation that captures the most important
effects of rotation onto the \lya\ line. 
We defer the reader to \citet{GaravitoCamargo2014} for complete
details on the explicit form of this approximation.
To fully explore the
parameter space using a Markov Chain Monte Carlo (MCMC) calculation with the
\texttt{emcee} Python library \citep{2013PASP..125..306F}. \texttt{emcee} 
is an open source optimized implementation of the affine-invariant 
MCMC sampler \citep{goodman2010ensemble}. 
The algorithm creates a number of walkers that,
during a sufficient number of steps, generate parameters' combinations
for a specific model.
For each timestep, the code calculates the likelihood of the
combination with respect to the observational data.
The walkers explore the parameter space sampling the Gaussian likelihood
function built as $\propto \exp(-\chi^2/2)$, where the $\chi^2$ follows
the definition in Equation \ref{eq:chi2}. 
We do not have a closed analytic expresion for $\hat{f}$, we compute it
by numerical integration of the analytical approximations found in
\citet{GaravitoCamargo2014}.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that separate two
classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dipersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya\ photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.6\textwidth]{vinf_cl.pdf}
  \includegraphics[width=0.6\textwidth]{P_cl.pdf}
  \includegraphics[width=0.6\textwidth]{sigma_cl.pdf}
\caption{Results from the multiphase  model.
  We only show results for the three parameters with a significant
  statistical difference between the models with the lowest $\chi^2/d.o.f.$
  values and its prior distribution. 
  These three parameters are the clump radial velocity at infinity 
  $v_{\infty, \rm{cl}}$; the probability that  the \lya\ photons were
  emitted in the clumps, $P_{\rm cl}$; the clump velocity dispersion,
  $\sigma_{\rm cl}$.
  The left column corresponds to the parameter's integrated distributions for
  models with the lowest $\chi^2/d.o.f.$ values (dotted line) compared against the
  parameter's prior integrated distributions (continuous line).
  The right column shows a scatter plot of $\chi^2/d.o.f.$
  and its corresponding physical parameters.
  The clump kinematic state is the key ingredient to
  reproduce \tol's \lya\ line morphology.
  \label{multiphaseresults}
}  
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{emcee_results.pdf}
\caption{Results from the Markov Chain Monte Carlo computation for
    the rotation model. 
    The gray scale indicates the point density in parameter space. 
    The dotted vertical lines in the histograms in the diagonal panels
    represent the 16th, 50th and 84th percentiles. 
    All parameters in the rotation model can be succesfuly constrained by \tol's
    \lya\ line morphology to the values written in the top region of
    the diagonal panels.
    \label{emceeresults}} 
\end{center}
\end{figure*}

@@

The fit to the observations is not perfect. 
However, in spite of the
simplicity of our models, this is the first time that the main
\tol\ features can be reproduced: a broad, highly symmetric,
single-peaked \lya\ line.

@@

In what follows we summarize the values of the kinematic parameters
that managed to explain \tol\'s \lya\ profile.

@@

With $14$ free parameters our first concern is discovering which are
the most relevant in reproducing \tol's line shape.
The results of the random forest classifier yield that the clump outflow velocity
$v_{\infty,\rm{cl}}$, the clump velocity dispersion $\sigma_{\rm cl}$ and the probability
that the \lya\ emission comes from the clumps $P_{\rm cl}$, are the
most influential parameters in finding a model with low
$\chi^2/d.o.f.$.

@@

In this paper we presented two kinematic models that independently
reproduce the so far unexplained observational features of \tol's
\lya\ emission line.  
One model is based on a multiphase ISM with random clump motions and
the other on gas bulk solid body rotation. 
It is the first time that an observed \lya\ profile can be fully
reproduced by either of these two kinematic conditions.
Our findings highlight the importance of including multiphase and/or
rotation conditions as kinematic features to model the \lya\ line.

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[An atypical \lya dwarf galaxy]{
Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy}
\author[J.E. Forero-Romero et al.]
{Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co},
Max Gronke$^2$, 
Maria Camila Remolina-Guti\'errez$^1$,
\newauthor
Nicol\'as Garavito-Camargo$^3$, 
Mark Dijkstra$^2$\\
%%
$^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
$^2$ Institute of Theoretical Astrophysics, University of Oslo,
Postboks 1029 Blindern, NO-0315 Oslo, Norway.\\
$^3$ Department of Astronomy, University of Arizona, 933 North Cherry
Avenue, Tucson, AZ 85721, USA. 
}

@@

Star-forming Compact Dwarf Galaxies (CDGs) 
resemble the expected pristine conditions of the first galaxies in the
Universe and are the best systems to test models on primordial galaxy
formation and evolution. 
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya emission line that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya spectrum can be explained
assuming either of these kinematic conditions.
The multiphase model requires a clump velocity dispersion of
at least \sigmaclump\ with outflows of \inftyclump\, while the
bulk rotation velocity is constrained to be $348^{+75}_{-48}$ \kms.
We argue that the multiphase model has more chances to be correct.
The clump velocity dispersion implies a dynamical mass of at least $2\times
10^{9}$ M$_{\odot}$, ten times its baryonic mass, which can be
explained if \tol\ is hosted by a dark matter halo of at least
$6\times 10^{11}$ M$_{\odot}$.
If future kinematic maps of \tol\ confirm the velocities suggested by the
multiphase model, it would provide additional support to expect this
condition to be present  in primordial \lya emitting galaxies.  
\end{abstract}

@@

Primordial galaxies have not been detected yet. 
However, dwarf star forming galaxies with a low metallicity content
are seen as templates to understand the early galaxy evolution process. 
Over fifty years ago it was realized that young galaxies could be
detected through a strong \lya line emission \citep{PartridgePeebles}.

This theoretical prediction was only confirmed thirty years later on
distant, relatively young, not primordial, galaxies \citep{1998ApJ...498L..93D}.
Currently Lyman Alpha Emitting (LAE) galaxies are commonly targeted
in surveys. 
The presence of the \lya emission line provides confirmation of
the distance to a galaxy while provides clues about the stellar
population and inter-stellar medium conditions regulating the
\lya emission.
A careful clustering analysis of LAEs can also yield clues about their link
to dark matter halos
\citep{2004AJ....128.2073H,2007ApJ...671..278G,2007ApJ...668...15K,2008MNRAS.391.1589O,2010MNRAS.409..184P,2013MNRAS.431.1777G,2016ApJ...828....5M}.

The \lya emission line is not exclusive of distant galaxies. 
There are local Universe surveys that target \lya emission in nearby
dwarf star forming galaxies.
The study of nearby LAE samples has allowed the study of other
indicators that might be more difficult to obtain for distant galaxies
such as galaxy morphology, dust attenuation, neutral hydrogen contents and
ionization state. See \cite{Hayes15} and references therein for a review.

However, the physical interpretation of \lya observations is
not straightforward \citep{LARS,2015ApJ...805...14R}. 
This is due to the resonant nature of the \lya line. 
A \lya photon follows a diffusion-like process before escaping
the galaxy or being absorbed by dust. 
The resulting line profile becomes sensitive to the dynamical, chemical
and thermal conditions in the interstellar medium. 
There are few analytical tools available to interpret the
emerging \lya line
\citep{Harrington73,1991ApJ...370L..85N,LoebRybicki,2006ApJ...645..792T}. 
They are applicable only in cases of highly symmetrical
conditions of the gaseous hydrogen distribution in the galaxy model,
which are hardly met in real astrophysical systems. 
For these reasons the interpretation of \lya observations
requires state-of-the-art Monte Carlo radiative transfer simulations.

Observed \lya line profiles usually present a single peak shifted
redwards from the line's center. 
Sometimes a double peak is present but the asymmetry persists with 
the peak on the red side being stronger. 
This can be explained as the results of multiple \lya photon
scatterings through an outflowing shell of neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

\tol\ is a compact star forming dwarf galaxy that presents a
strong \lya emission with puzzling 
features: the line is highly symmetric, single peaked and broad 
\citep{Thuan97}.
This kind of profile also raises the question whether some high
redshift LAEs have asymmetric lines because the blue half was
truncated by the intergalactic medium \citep{2007MNRAS.377.1175D}. 
In this case the \lya radiation could emerge as a low surface
brightness glow, which may be connected to \lya halos, while also
influencing the way LAEs can be used as a probe of reionization,
see the review by \cite{2014PASA...31...40D} and the references
therein.

Attempts to explain the atypical \lya features in \tol\ with
conventional models based on a expanding shell have not been succesful
so far \citep{mashesse03,2015A&A...578A...7V}. 
This justifies a new study using two new physically motivated
situations: a multiphase outflowing interstellar medium and bulk rotation for a
galaxy without outflows.

The main motivation for the outflowing multiphase model (as presented
in \cite{Gronke2016}) is that some dwarf galaxies are expected to
present outflows.  
Observationally, outflows have been detected in few local dwarf galaxies
\citep{1998ApJ...506..222M,2005MNRAS.358.1453O}. 
Besides, clumpy multiphase outflows are capable to explain \lya features around
star-forming galaxies
\citep{2010ApJ...717..289S,2012MNRAS.424.1672D}.

@@

The models we use correspond to simplified geometrical configurations.
This allows us to perform a deep exploration of parameter space and
gain some physical insight into \tol's kinematic properties.
In this paper we show, for the first time in the literature, that \tol's
\lya profile can be explicitly modeled be either of these two
models.

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be sucessful.
We focus on presenting two simple and succesful models that
manage, for the first time, to account for \tol's \lya line main
features.

In the next section We review the observational characteristics of
\tol, then we summarize the main features in the multiphase and
rotation models to explain how we fit their free parameters 
to the \tol's \lya line shape.
We use the results to interpret them in terms of the galaxy's
dynamical mass and argue why the multiphase model should be prefered
over the rotation model.
We close by discussing the additional observational information needed
to clarify the kinematic nature of \tol.

@@

\begin{table*}
  \centering
\begin{tabular}{l|lrrl}
\hline
  Parameter  & Description & Fiducial value & Allowed range & Units \\ \hline\hline
$v_{\infty,\,{\rm cl}}$ & Radial cloud velocity & $100.0$ & [$0.0$, $800.0$] & $\,{\rm km\,s}^{-1}$ \\
$\sigma_{\rm {\rm cl}}$ & Random cloud motion &$40.0$ & [$5.0$, $100.0$] & $\,{\rm km\,s}^{-1}$ \\
$P_{\rm {\rm cl}}$ & Probability to be emitted in cloud & $0.35$ & [$0.0$, $1.0$] &  \\
$r_{\rm {\rm cl}}$ & Cloud radius & $100.0$ & [$30.0$, $200.0$] & $\,{\rm pc}$ \\
$H_{\rm em}$ & Emission scale radius & $1000.0$ & [$500.0$, $3.0\times 10^{3}$] & $\,{\rm pc}$ \\
$f_{\rm {\rm cl}}$ & Cloud covering factor & $3.5$ & [$0.8$, $8.0$] &  \\
$T_{\rm {\rm ICM}}$${}^{\dagger}$ & Temperature of ICM & $10^{6}$ & [$3.0\times 10^{5}$, $5.0\times 10^{7}$] & $\,{\rm K}$ \\
$n_{\rm HI,\,{\rm ICM}}$${}^{\dagger}$ & \HI number density in ICM & $5.0\times 10^{-8}$ & [$10^{-12}$, $10^{-6}$] & $\,{\rm cm}^{-3}$ \\
$\sigma_{\rm i}$ & Width of emission profile & $50.0$ & [$5.0$, $100.0$] & $\,{\rm km\,s}^{-1}$ \\
$T_{\rm {\rm cl}}$${}^{\dagger}$ & Temperature in clouds & $10^{4}$ & [$5.0\times 10^{3}$, $5.0\times 10^{4}$] & $\,{\rm K}$ \\
$\beta_{\rm cl}$ & Steepness of the radial velocity profile & $1.5$ & [$1.1$, $2.5$] &  \\
$\tilde\sigma_{\rm d, cl}$${}^{\dagger}$ & Dust content in clumps &
  $3.2\times10^{-22}$ & [$4.7\times 10^{-24}$, $1.6\times 10^{-21}$] &
  $\,\cm^{2}$ \\
$\zeta_d$${}^{\dagger}$ & Ratio of ICM to cloud dust abundance &
  $0.01$ & [$10^{-4}$, $0.1$] & \\
$n_{\rm HI,\,{\rm cl}}$${}^{\dagger}$ & \HI number density in
  clouds & $0.35$ & [$0.03$, $3.0$] & $\,{\rm cm}^{-3}$ \\
\hline
\end{tabular}
  \caption{Overview of the parameters in the multiphase model and its fiducial
    values. Variables marked with ${}^{\dagger}$ were drawn in
    log-space. Table reproduced from \citet{Gronke2016}.}
  \label{tab:models}
\end{table*}

@@

Archival Chandra X-ray data \footnote{\url{http://cxc.harvard.edu/cda/}} does not show any
detection for \tol. 
This lack of detection motivates our assumption that \lya emission is
powered by star formation only.

The observed flux for the \lya line is $\sim
8.1\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \citep{Thuan97}.
The \lya Equivalent Width is $70$\AA\ and its H$\beta$ flux is 
$1.62\times 10^{-14}$ erg cm$^{-2}$ s$^{-1}$ \AA${-1}$
which gives a \lya/H$\beta$ flux ratio of
4.9$\pm$0.1 \citep{Izotov04}.
Comparing the \lya/H$\beta$ ratio with the theoretical
expectation from case B recombination of $23.3$ \citep{Hummer1987} one
can estimate an escape fraction of $20$\% for \lya radiation.

The \lya flux values correspond to a luminosity of
$L_{Ly\alpha}=2.2\times 10^{42}$ erg s$^{-1}$, which in turn
translates  into a lower bound for the star formation rate of $2.0$
M$_{\odot}$ yr$^{-1}$ after using a standard conversion factor between
luminosity and star formation rate of $9.1\times 10^{-43}$
$L_{Ly\alpha}$ M$_{\odot}$ yr$^{-1}$ 
without any
correction by extinction
\citep{Kennicutt98}.

@@

We compute the projected half-light radius to be $R_s=1.5\pm0.1$ kpc 
from the surface brightness profiles reported by \cite{2003A&A...410..481N}. 
Assuming spherical geometry, one can translate this value into a 3D
half-light radius of $r_s=3R_s/2=2.25$ kpc.
Imaging observations by \cite{2001AJ....121..169F} show that \tol\ is
an isolated field galaxy not belonging to a group or cluster.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.55\textwidth]{CLARA-TOL-main.pdf}
\caption{Broad, single peaked and highly symmetric \lya emission of \tol.
  Dots correspond to the observational data \citep{mashesse03}. 
The thick red and thin blue curves represent our best fit
to the data using the full radiative transfer simulation with a
rotation and multiphase model, respectively. 
These two different models are able to reproduce the main morphological
features of \tol's \lya line emission.\label{fig:results}}
\end{center}
\end{figure*}

@@

The rotation model corresponds to the work presented in
\citep{GaravitoCamargo2014} based on the Monte Carlo code
\texttt{CLARA} \citep{CLARA}. 
In that model the \lya photons are propagated 
within a spherical and homogeneous cloud of HI gas undergoing solid
body rotation.
The sphere is fully characterized by three parameters: the HI line's
center optical  depth $\tau$ measured from the center to its surface, the HI
temperature $T$, and the linear surface velocity $V_{\rm max}$.  
The observed line profile also depends on $\theta$,  the angle between the plane
perpendicular to the rotation axis and the observational
line-of-sight.  
In this model, dust only changes the overall line
normalization and only weakly its shape, i.e. 
dust cannot change the line symmetry or induce a change in the number
of line peaks, moreover it does not change the line width by more than
$1\%$ ($5$ \kms\ in the case of \tol), an effect too small to be
observed at the resolution at which we have \tol's line profile and
also negligeable compared to the influence of the other free parameters in the model.
For these reasons we do not include any dust model.

We use an analytical approximation that captures the most important
effects of rotation onto the \lya line. 
We defer the reader to \citet{GaravitoCamargo2014} for complete
details on the explicit form of this approximation.
To fully explore the
parameter space using a Markov Chain Monte Carlo (MCMC) calculation with the
\texttt{emcee} Python library \citep{2013PASP..125..306F}. \texttt{emcee} 
is an open source optimized implementation of the affine-invariant 
MCMC sampler \citep{goodman2010ensemble}. 
The algorithm creates a number of walkers that,
during a sufficient number of steps, generate parameters' combinations
for a specific model.
For each timestep, the code calculates the likelihood of the
combination with respect to the observational data.
The walkers explore the parameter space sampling the Gaussian likelihood
function built as $\propto \exp(-\chi^2/2)$, where the $\chi^2$ follows
the definition in Equation \ref{eq:chi2}. 
We do not have a closed analytic expresion for $\hat{f}$, we compute it
by numerical integration of the analytical approximations found in
\citet{GaravitoCamargo2014}.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that separate two
classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dipersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.6\textwidth]{vinf_cl.pdf}
  \includegraphics[width=0.6\textwidth]{P_cl.pdf}
  \includegraphics[width=0.6\textwidth]{sigma_cl.pdf}
\caption{Results from the multiphase  model.
  We only show results for the three parameters with a significant
  statistical difference between the models with the lowest $\chi^2/d.o.f.$
  values and its prior distribution. 
  These three parameters are the clump radial velocity at infinity 
  $v_{\infty, \rm{cl}}$; the probability that  the \lya photons were
  emitted in the clumps, $P_{\rm cl}$; the clump velocity dispersion,
  $\sigma_{\rm cl}$.
  The left column corresponds to the parameter's integrated distributions for
  models with the lowest $\chi^2/d.o.f.$ values (dotted line) compared against the
  parameter's prior integrated distributions (continuous line).
  The right column shows a scatter plot of $\chi^2/d.o.f.$
  and its corresponding physical parameters.
  The clump kinematic state is the key ingredient to
  reproduce \tol's \lya line morphology.
  \label{multiphaseresults}
}  
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{emcee_results.pdf}
\caption{Results from the Markov Chain Monte Carlo computation for
    the rotation model. 
    The gray scale indicates the point density in parameter space. 
    The dotted vertical lines in the histograms in the diagonal panels
    represent the 16th, 50th and 84th percentiles. 
    All parameters in the rotation model can be succesfuly constrained by \tol's
    \lya line morphology to the values written in the top region of
    the diagonal panels.
    \label{emceeresults}} 
\end{center}
\end{figure*}

@@

The fit to the observations is not perfect. 
However, in spite of the
simplicity of our models, this is the first time that the main
\tol\ features can be reproduced: a broad, highly symmetric,
single-peaked \lya line.

@@

In what follows we summarize the values of the kinematic parameters
that managed to explain \tol\'s \lya profile.

@@

With $14$ free parameters our first concern is discovering which are
the most relevant in reproducing \tol's line shape.
The results of the random forest classifier yield that the clump outflow velocity
$v_{\infty,\rm{cl}}$, the clump velocity dispersion $\sigma_{\rm cl}$ and the probability
that the \lya emission comes from the clumps $P_{\rm cl}$, are the
most influential parameters in finding a model with low
$\chi^2/d.o.f.$.

@@

In this paper we presented two kinematic models that independently
reproduce the so far unexplained observational features of \tol's
\lya emission line.  
One model is based on a multiphase ISM with random clump motions and
the other on gas bulk solid body rotation. 
It is the first time that an observed \lya profile can be fully
reproduced by either of these two kinematic conditions.
Our findings highlight the importance of including multiphase and/or
rotation conditions as kinematic features to model the \lya line.
